Hydras and Cie (Draft) \\ Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq

### 2021-09-19 doc/JFLA2022/paper.tex
minor corrections

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 
Laurie Kirby and Jeff Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}.
This article describes a game between two players: Hercules and a hydra.
A short description  can be found in Andrej Bauer's blogpost~\cite{bauer2008} as well as a simulator~\cite{BauerHydra}.
In a few words:
\begin{itemize}
\item A hydra is a finite tree, traditionnaly presented with the root at the bottom, the leaves of which are called \emph{heads}
  (Fig~\vref{fig:iota-example}).
\item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
  then some sub-tree $h$ of the hydra is copied a certain amount $i$ of times. The number $i$ of copies and the copied sub-hydra $h$ may depend of the considered variant of the game
  or the time elapsed since the beginning of the fight.
\end{itemize}

@@

The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-community} contains a development by José Grimm~\cite{grimm:hal-00911710}. This library -- dedicated  to the implementaion in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics --- contains a generalization of two types of ordinals (\texttt{T1} for the ordinal $\epsilon_0$ and
\texttt{T2} for $\Gamma_0$) to a larger ordinal defined by Ackermann, and relates these data structures with Bourbaki's set theory.
It is interesting to note that both libraries \textit{Gaia} and \textit{hydra-battles} contain their own (isomorphic) definitions of
\texttt{T1} and \texttt{T2}, originated in~\cite{CantorContrib}.
The types --- defined in different modules --- cannot be considered as equal. Even more, some operations like addition, multiplication, etc. have now different definitions.

It would be interesting to build a library which could be used as a gate giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}.
For instance, the associativity of multiplication of ordinals below $\epsilon_0$, already proven in \textit{Gaia} was tranferred ``for free'' to the data structures of \textit{hydra-battles}, thanks to the isomorphism between both implementations.

>>>>

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 
Laurie Kirby and Jeff Paris: \emph{Accessible Independence Results for Peano Arithmetic}~\cite{KP82}.
This article describes a game between two players: Hercules and a hydra.
A short description  can be found in Andrej Bauer's blogpost~\cite{bauer2008} as well as a simulator~\cite{BauerHydra}.
In a few words:
\begin{itemize}
\item A hydra is a finite tree, traditionnaly presented with the root at the bottom, the leaves of which are called \emph{heads}
  (Fig~\vref{fig:iota-example}).
\item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
  then some sub-tree $h$ of the hydra is copied a certain amount $i$ of times. The number $i$ of copies and the copied sub-hydra $h$ may depend of the considered variant of the game
  or the time elapsed since the beginning of the fight.
\end{itemize}

@@

The Gaia project~\cite{Gaia}, also maintained on \textit{Coq-Community} contains a development by José Grimm~\cite{grimm:hal-00911710}. This library -- dedicated  to the implementaion in \textit{Coq (SSreflect)} of books from  Bourbaki's Elements of Mathematics --- contains a generalization of two types of ordinals (\texttt{T1} for the ordinal $\epsilon_0$ and
\texttt{T2} for $\Gamma_0$) to a larger ordinal defined by Ackermann, and relates these data structures with Bourbaki's set theory.
It is interesting to note that both libraries \textit{Gaia} and \textit{hydra-battles} contain their own (isomorphic) definitions of
\texttt{T1} and \texttt{T2}, originated in~\cite{CantorContrib}.
The types --- defined in different modules --- cannot be considered as equal. Even more, some operations like addition, multiplication, etc. have now different definitions.

It would be interesting to build a ``gate'' giving access to both the combinatorial results of \textit{hydra-battles} and the set-theoretic content of \textit{Gaia}.
For instance, the associativity of multiplication of ordinals below $\epsilon_0$, already proven in \textit{Gaia} was tranferred ``for free'' to the data structures of \textit{hydra-battles}, thanks to the isomorphism between both implementations.

Another direction would be to write in \textit{Coq} a formal proof of the statement of the second theorem of Kirby and Paris, using the formalization of Peano Arithmetic by O'Connor~\cite{Goedel}.
